---
title: "Erindi (talks)"
author: "Valtýr"
date: "09/12/2019"
---
<!-- `r anicon::faa("gamepad", animate='ring')` --> 
## Sources and Pathways of Microplastics to Icelandic Seas (**a talk**)
### 29. nóvember 2019

This is a talk about the results of a report that I, and a colleague, made for the Environmental Ministry of Iceland. Below is the talk <i class="fas fa-chalkboard-teacher"></i> and [here is the report <i class="fas fa-book"></i>](https://harkanatta.github.io/orplastskyrslaUAR/) (*in Icelandic*)

```{r echo=FALSE, fig.cap="Click the slideshow above, press `'c'` to clone it (opens on a seperate window) and then press `'f'` for fullscreen."}
knitr::include_url("https://harkanatta.github.io/mp-mirror")
```

## Örplast í hafinu við Ísland - Helstu uppsprettur, magn og farleiðir í hafið
### 31. október 2019

Létt erindi um niðurstöður skýrslu sem ber sama heiti og má finna [hér <i class="fas fa-book"></i>](https://harkanatta.github.io/orplastskyrslaUAR/). Erindið má sjá [hér <i class="fas fa-chalkboard-teacher"></i>](https://harkanatta.github.io/innragrunn)

```{r echo=FALSE, fig.cap="Click the slideshow above, press `'c'` to clone it (opens on a seperate window) and then press `'f'` for fullscreen.", include=FALSE}
knitr::include_url("https://harkanatta.github.io/innragrunn")
```

## The effects of a herring mass death on the bottom biota
### October 2014 and November 2015

[The effects of a herring mass death on the bottom biota in Kolgrafafjörður Iceland <i class="fas fa-chalkboard-teacher"></i>](http://www.vegagerdin.is/vefur2.nsf/c16a923a6485025d0025793b003db0e2/b81be193196d902500257d85004e57c2/$FILE/1630%20ValtyrSigurdsson1630.pdf) (*Masters thesis*). This talk was held on an event in 2014 by The Icelandic Road and Coastal Administration (which funded the project) and again, slightly modified, at the [Icelandic Biology Conference in 2015](https://biologia.is/liffraediradstefnan-2015/dagskra/)


