---
title: Kelp
author: Valtýr
date: '2019-10-24'
slug: þari
categories:
  - R
tags:
  - projects
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/iconid.png
---

In summertime we marine biologist get to venture out in the sun. Despite regular winds most of the year, summers here are great. But at these high latitudes they are short, only about 3 months. Unfortunately for me, I can get `r anicon::nia("sea sick",animate="float",speed="slow")` but still I manage to work my way around it.


```{r tafla, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, out.width='70%'}
#https://stackoverflow.com/questions/1395528/scraping-html-tables-into-r-data-frames-using-the-xml-package
library(XML)
library(RCurl)
library(rlist)
library(DT)
theurl <- getURL("http://www.ssnv.is/is/uppbyggingarsjodur/styrkhafar",.opts = list(ssl.verifypeer = FALSE) )
tables <- htmlParse(theurl, encoding = 'UTF-8')
tables <- readHTMLTable(tables)
tables <- list.clean(tables, fun = is.null, recursive = FALSE)
n.rows <- unlist(lapply(tables, function(t) dim(t)[1]))
tafla <- tables[[which.max(n.rows)]]
tafla[,3] <- gsub("\\.", "", tafla[,3])
tafla[,3] <- as.numeric(tafla[,3])/1000000
colnames(tafla) <- c("Styrkhafi","Verkefni","Milljónir")

widgetframe::frameWidget(datatable(tafla))
```


```{r echo=FALSE, fig.cap="Click the slideshow above, press `'c'` to clone it (opens on a seperate window) and then press `'f'` for fullscreen."}
knitr::include_url("https://harkanatta.github.io/kajakslides")
```

Last summer we wanted to estimate how extensive our kelp beds are in Húnaflói to see if there was some extra to be harvested for various purposes. We started off with looking out for sites of interest on kayaks (also took advice from local fishermen) then followed up with an [underwater drone](https://www.sofarocean.com/products/trident). The video recordings from the drone were used for classifying the kelp and other interesting things down there.

```{r yfirlitskort, fig.cap="Click on the points to see figures from the area (then right-click the image and select an option to view it better).", echo=FALSE, error=FALSE, fig.align="center", message=FALSE, warning=FALSE, out.width='100%'}

stodvar <- sf::read_sf(dsn="skjol/stodvar.shp")
#dypi20 <- rgdal::readOGR(dsn="skjol/20mDypi.shp", verbose = FALSE)
dypi10 <- rgdal::readOGR(dsn="skjol/10mlengjur.shp", verbose = FALSE)
dyptarl <- rgdal::readOGR(dsn="skjol/dyptarl.shp", verbose = FALSE)

library(raster)
skogar <- shapefile("skjol/Konnun.shp")
skogar@data$SV[c(15:16,19,21:23)] <- c("Hamarsbúð",    "Hindisvík",     "Finnsstaðanes",  "Bjargarbjörg",   "Kálfshamarsvík", "Sölvabakki")
skogar <- skogar[skogar@data$STIG==3,]

#,vefur=rep("https://harkanatta.github.io/orplast2019/slaedur/slaedur.html#1",length(skogar@data$SV)))
a <- list()
b <- list()
for (i in 1:length(dypi10@data[,1])) {
 a[i] <- round(area(dypi10[i,1])/1000000)
 b[i] <- as.character(dypi10[i,]$strond)
}
DF2 <- data.frame(Stadur=unlist(b),Km²=unlist(a),Dypi=rep("10 m",length(dypi10@data[,1])))
DF <- data.frame(Stadur=unlist(b),Km2=unlist(a),Dypi=rep("10 m",length(skogar@data$SV)))
library(plyr)
DF <- ddply(DF,("Stadur"),summarize,km²=sum(Km2))

images = c("https://raw.githubusercontent.com/harkanatta/ssnv_trident/master/skjol/skogar/sk6/hafnir.jpeg",
           "https://raw.githubusercontent.com/harkanatta/ssnv_trident/master/skjol/skogar/sk4/samsett.jpeg",
           "https://raw.githubusercontent.com/harkanatta/ssnv_trident/master/skjol/skogar/Finnssta%C3%B0anes/hrossa.jpeg",
           "https://raw.githubusercontent.com/harkanatta/ssnv_trident/master/skjol/skogar/sk1/stodvar/st1/sk1-st1B.jpg",
           "https://raw.githubusercontent.com/harkanatta/ssnv_trident/master/skjol/skogar/sk2/samsett.jpeg",
           "https://raw.githubusercontent.com/harkanatta/ssnv_trident/master/skjol/skogar/sk5/stodvar/st1/sk5-st1.jpg",
           "https://raw.githubusercontent.com/harkanatta/ssnv_trident/master/skjol/skogar/sk3/samsett.jpeg")

urls=c("https://harkanatta.github.io/orplast2019/slaedur/slaedur.html#1","https://harkanatta.github.io/orplast2019/slaedur/slaedur.html#3","https://harkanatta.github.io/orplast2019/slaedur/slaedur.html#18","https://harkanatta.github.io/orplast2019/slaedur/slaedur.html#21","https://harkanatta.github.io/orplast2019/slaedur/slaedur.html#11","https://harkanatta.github.io/orplast2019/slaedur/slaedur.html#20","https://harkanatta.github.io/orplast2019/slaedur/slaedur.html#7")

litir <- RColorBrewer::brewer.pal(7,"Accent")
library(mapview)
library(magrittr)
map <- mapview(dypi10,col.regions="#cb5600",map.types="Stamen.TerrainBackground", legend = FALSE,popup=leafpop::popupTable(DF2))+mapview(stodvar, popup = leafpop::popupImage(images,width="500"),legend=FALSE,col.regions=litir,cex=14)+mapview(skogar,popup=leafpop::popupTable(DF),label=skogar@data$SV,legend=FALSE)+mapview(dyptarl,lwd=0.3,legend=FALSE)
map

# library(magrittr)
# img <- "https://github.com/harkanatta/ssnv_trident/blob/master/graphs/tvologo.jpg?raw=true"
# map %>% leafem::addLogo(img, width = '20', height = '25',offset.y = 20,offset.x = 80,alpha = 0.7) %>% leaflet.extras::addFullscreenControl(pseudoFullscreen = T)

```

Click the points on the map above to see figures from the area (then right-click the image and select an option to view it better).
